---
title: "lobster-project"
author: "Anna Calle"
date: "11/12/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include = FALSE}
# Load packages and read .csv files
#install.packages("vcdExtra")
library(tidyverse)
library(vcdExtra)
library(car)
library(onewaytests)
library(kableExtra)
library(effsize)
library(extrafont) # Easier font changes
#font_import()
# loadfonts()

library(FSA) # for dunnTest function
library(ggsignif) # To add significance brackets

size_abundance <- read_csv("lobster_size_abundance.csv")
traps <- read_csv("lobster_traps.csv")

#You will be describing lobster size, abundance and fishing pressure at five Long-Term Ecological Research (LTER) Sites in the Santa Barbara Channel close to the mainland: Arroyo Quemado (AQUE), Naples Reef (NAPL), Mohawk Reef (MOHK), Isla Vista (IVEE), Carpinteria (CARP).
```

### 1. Lobster abundance and fishing pressure (2012 - 2017)
Describe trends in lobster abundance (counts) and fishing pressure (trap buoys) at the five locations from 2012 - 2017. Ignore transect information - we are only interested in evaluating abundance and pressure on the
order of SITE. Note: you are not expected to use regression here - just think of ways to clearly describe annual totals visually and in text, noting important trends, events and differences.

```{r, message = FALSE}
# Transform data into case form

size_abundance <- data.frame(size_abundance)
size_abun_case <- expand.dft(size_abundance, freq = "COUNT")

#create table and graph of lobster abundance by site and year

abundance_table <- size_abun_case %>% 
  count(YEAR, SITE) %>% 
  spread(SITE, n)

#summarized abundances 
abundance_summary <- size_abundance %>% 
  group_by(SITE, YEAR) %>% 
  summarize(abundance = sum(COUNT),
            sd = sd(COUNT))

abundance_summary

#Figure of abundances
abundance_figure <- ggplot(abundance_summary, aes(x = YEAR, y = abundance)) +
  geom_line(aes(linetype = SITE)) +
  theme_classic()+
  theme(axis.text.x=element_text(angle=90,hjust=1))+
  labs(y = "Abundance of Lobsters", x= "Year") +
  #geom_errorbar(aes(ymin=abundance - sd ,ymax= abundance + sd ), width=.2) +
  theme(axis.text=element_text(size=11))+
  theme(axis.title=element_text(size=12))+
  scale_y_continuous(expand = c(0,0), limits = c(0,750))+
  scale_x_continuous(expand = c(0,0), limits = c(2012,2017)) +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank(), axis.line = element_line(colour = "black"),text = element_text(family = "Times New Roman"))

abundance_figure


#Final table but not going to incorperate in final paper
final_table_lobsters <- kable(abundance_table, col.names = c("Year", "Arroyo Quemado", "Carpinteria", "Isla Vista", "Mohawk Reef", "Naples Reef"), align = "c") %>% 
   kable_styling(bootstrap_options = "striped", full_width = F) %>% 
  kable_styling(bootstrap_options = "hover") %>% 
  column_spec(1, bold = T) %>% 
  row_spec(0, bold = T, color = "black")

final_table_lobsters #finalized table


  
```

__Figure 1. Abundance of Lobsters in Santa Barbara Channel (2012-2017).__ Lobster counts measured at Long Term Ecological Research Sites (LTERS) in the Santa Barbara Channel: Arroyo Quemado (AQUE), Naples Reef (NAPL), Mohawk Reef (MOHK), Isla Vista (IVEE), Carpinteria (CARP) from 2012 to 2017. Error bars indicate ± 1 standard deviation. 


```{r, message = FALSE}
#create table and graph of trap buoy counts by site and year 

traps <- data.frame(traps)

traps_case <- traps %>%   
  filter(SITE != "ABUR", SITE != "AHND", SITE != "AHND to AQUE", SITE != "GOLB") %>% 
  group_by(SITE, YEAR) %>% 
  summarize(trap_count = sum(TRAPS))

#figure of graph 

traps_figure <- ggplot(traps_case, aes(x = YEAR, y = trap_count)) +
  geom_line(aes(linetype = SITE)) +
  theme_classic() +
  theme(axis.text.x=element_text(angle=90,hjust=1)) +
  labs(y = "Trap Count", x= "Year") +
   theme(axis.text=element_text(size=11))+
  theme(axis.title=element_text(size=12))+
  scale_y_continuous(expand = c(0,0), limits = c(0,1200))+
  scale_x_continuous(expand = c(0,0), limits = c(2012,2017)) +
   theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank(), axis.line = element_line(colour = "black"),text = element_text(family = "Times New Roman"))

traps_figure

```

__Figure 2. Fishing Pressure in Santa Barbara Channel (2012-2017).__ Trap Buoy counts measured at Long Term Ecological Research Sites (LTERS) in the Santa Barbara Channel: Arroyo Quemado (AQUE), Naples Reef (NAPL), Mohawk Reef (MOHK), Isla Vista (IVEE), Carpinteria (CARP) from 2012 to 2017. Error bars indicate ± 1 standard deviation. No trap buoys were measured for Isla Vista or Naples Reef. 

While all five sites had higher lobster counts in 2017 than 2012, Isla Vista and Carpinteria both experienced a much larger increase in lobster abundance (figure 1). In 2017, Carpinteria had a recorded lobster count of 705, compared with 98 in 2012. In contrast, Arroyo Quemado had an increase of only 29 over the same time period. Figure 2 shows fishing pressure in the Santa Barbara Channel, measured by the number of trap buoys recorded at each site. As can be seen from the data, fishing pressure decreased at all sites between 2012 and 2017. (no trap buoys were recorded at the two MPA sites, Isla Vista and Naples Reef). The decrease in fishing pressure is a possible explanation for the increase in lobster counts in the Santa Barbara Channel between 2012 and 2017. Interestingly, lobster counts at the MPA sites both increased even though fishing pressure was not a factor. Lobster counts at Isla Vista, for example, increased by 580, meaning that there are likely other factors contributing to the increase in lobster abundance. 

### 2. Compare mean lobster size by site in 2017
Compare mean lobster sizes (carapace length (mm)) across the five sites for lobster observations collected in 2017. Warning: the size data are not in tidy format. There are rows that contain size information for multiple
lobsters observed (e.g., if the researcher saw 3 lobsters all with carapace length ~ 60 mm, then they will have a single row where COUNT = 3 and SIZE = 60). You’ll want to get this into case format - where each lobster has its own row - before doing statistical analyses. There are many ways to do this. One hint: function expand.dft in the vcdExtra package (it doesn’t like tibbles, so you might need to coerce to data.frame first).
```{r}
# Compare mean lobster sizes across sites 

#Create dataframe with only 2017 observations to run ANOVA

mean_sizes <- size_abun_case %>% 
  filter(YEAR == 2017) 

#Use Levene's test for equal variances 

#H0: Variances are equal
#HA: Variances are not equal 

lobster_variance <- leveneTest(SIZE ~ SITE, data = mean_sizes)

lobster_variance

#One way ANOVA for parametic data, because variances are not equal

#H0: There is no significant difference in mean lobster sizes across different sites
#HA: At least two sites have signficantly different means


lobster_aov <- aov(SIZE ~SITE, data = mean_sizes)
summary(lobster_aov)


# Kruskal-Wallis reveals that there are at least two means that are significantly different, therefore, we should run Tukey's HSD to explore further (post hoc test)

lobster_ph <- TukeyHSD(lobster_aov)

lobster_ph


# Mean lobster sizes differ significantly between Naples Reef (NAPL) and Carpinteria (CARP) (p = 0.023), Naples Reef (NAPL) and Isla Vista (IVEE) (0.004), Naples Reef (NAPL) and Mohawk Reef (MOHK) (p = 0.057).     


sizes_2017 <- size_abun_case %>% 
  filter(YEAR == 2017) %>% 
  group_by(SITE) %>% 
  summarize(
    mean_size=mean(SIZE),
    sd_size = sd(SIZE)
  )


figure <- ggplot(sizes_2017, aes(x = SITE, y = mean_size)) +
  geom_col(colour = NA, fill = "gray50", width = 0.5) +
  geom_errorbar(aes(ymax = mean_size + sd_size, ymin = mean_size - sd_size), width = 0.1) +
  scale_y_continuous(expand = c(0,0), limits = c(0,95)) +
  labs(y= "Mean Lobster Carapase Size (mm)") +
  scale_x_discrete(labels = c("AQUE","CARP","IVEE","MOHK", "NAPL")) +
  xlab("\nLocation") +
  annotate("text", x = 1, y = 88.5, label = "a,b", family = "Times New Roman") +
  annotate("text", x = 2, y = 88, label = "a", family = "Times New Roman") +
  annotate("text", x = 3, y = 88, label = "a", family = "Times New Roman") +
  annotate("text", x = 4, y = 84, label = "a,b", family = "Times New Roman") +
  annotate("text", x = 5, y = 90, label = "b", family = "Times New Roman") +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank(), axis.line = element_line(colour = "black"),text = element_text(family = "Times New Roman"))
  
  figure
```
__Figure 3. Mean lobster sizes in the Santa Barbara Channel (2017).__ Mean lobster carapace sizes (mm) measured at Long Term Ecological Research Sites (LTERS) in the Santa Barbara Channel: Arroyo Quemado (AQUE), Naples Reef (NAPL), Mohawk Reef (MOHK), Isla Vista (IVEE), Carpinteria (CARP) in 2017. Error bars indicate ± 1 standard deviation. Like letters indicate values that do not differ significantly (by one-way ANOVA with Tukey’s HSD; F(4,2355) = 3.24, p = 0.008), α = 0.05 for all post-hoc pairwise comparisons)


Lobster sizes (mm) differed signficantly between sites in 2017 (one-way ANOVA,F(4,2355) = 3.24, p = 0.008). Post-hoc analysis by Tukey’s HSD revealed that of the five sites studied, lobster sizes Naples Reef (n = 112) differed significantly from lobster sizes at Carpinteria (n = 705) and Isla Vista (n = 606) [pairwise p < 0.05, $\alpha$ = 0.05]. Mean lobster size at Naples Reef was the largest at 76.23 mm, compared with the other locations, all of which had means between 71 and 73 mm. Arroyo Quemado, the only location that did not differ signficantly from any, had a mean size of 73.89 mm. 



### 3. Changes in lobster size at MPA and non-MPA sites (comparing only 2012 and 2017 sizes)
At Isla Vista and Naples Reef, the two protected MPA sites (with zero fishing pressure), how do lobster sizes in 2012 and 2017 compare? At the non-MPA sites?
```{r}

#two years (2012 and 2017)
#two managment levels 

#5 separate t-tests (unpaired) between 2012 and 2017 at each site. No t tests between sites. That will be in our analysis 


########First look at the three non MPA sites######################

#Site 1: AQUE (Arroyo Quemado)

AQUE_2017 <- size_abun_case %>% 
  filter(YEAR == 2017, SITE == "AQUE")
mean(AQUE_2017$SIZE)

AQUE_2012 <- size_abun_case %>% 
  filter(YEAR == 2012, SITE == "AQUE")

#ftest to compare variances at AQUE in 2012 and 2017 
# null: Variances are equal (ratio is equal to 1)
# alternative: Variances are not equal (ratio is not equal to 1)

AQUE_ftest <- var.test(AQUE_2012$SIZE, AQUE_2017$SIZE)

AQUE_ftest

#ftest indicates that variances are equal (p =0.29). Use students t-test to compare sizes at AQUE in 2012 and 2017

AQUE_ttest <- t.test(AQUE_2012$SIZE, AQUE_2017$SIZE, var.equal = TRUE)
  
AQUE_ttest

#Our two sample t-test indicates that there is no significant difference in lobster size at Arroyo Quemado (p=0.209) in 2012 (mean = 71) and 2017 (mean = 73.89) 


#Site 2: MOHK (Mohawk Reef)

MOHK_2017 <- size_abun_case %>% 
  filter(YEAR == 2017, SITE == "MOHK")

mean(MOHK_2017$SIZE)

MOHK_2012 <- size_abun_case %>% 
  filter(YEAR == 2012, SITE == "MOHK")

#ftest to compare variances in lobster size at NAPL in 2012 and 2017 
# null: Variances are equal (ratio is equal to 1)
# alternative: Variances are not equal (ratio is not equal to 1)

MOHK_ftest <- var.test(MOHK_2012$SIZE, MOHK_2017$SIZE)

MOHK_ftest

#ftest indicates that variances are equal (p =0.15). Use students t-test to compare sizes at NAPL in 2012 and 2017

MOHK_ttest <- t.test(MOHK_2012$SIZE, MOHK_2017$SIZE, var.equal = TRUE)
  
MOHK_ttest

#Our two sample t-test indicates that there is a significant difference in lobster size at Mohwak Reef (p=0.000067) in 2012 (mean = 77) and 2017 (mean = 72) 

# Test of effect size 

effsize <- cohen.d(MOHK_2012$SIZE, MOHK_2017$SIZE)

effsize

# effect size was medium. 0.5408116 

#Site 3: CARP (Carpinteria)

CARP_2017 <- size_abun_case %>% 
  filter(YEAR == 2017, SITE == "CARP")

mean(CARP_2017$SIZE)

CARP_2012 <- size_abun_case %>% 
  filter(YEAR == 2012, SITE == "CARP")

#ftest to compare variances in lobster size at NAPL in 2012 and 2017 
# null: Variances are equal (ratio is equal to 1)
# alternative: Variances are not equal (ratio is not equal to 1)

CARP_ftest <- var.test(CARP_2012$SIZE, CARP_2017$SIZE)

CARP_ftest

#ftest indicates that variances are equal (p =0.204). Use students t-test to compare sizes at NAPL in 2012 and 2017

CARP_ttest <- t.test(CARP_2012$SIZE, CARP_2017$SIZE, var.equal = TRUE)
  
CARP_ttest

#Our two sample t-test indicates that there is no significant difference in lobster size at Isla Vista (p=0.182) in 2012 (mean = 74.35) and 2017 (mean = 72.29) 


#################MPA sites######################

#Site 4: Isla Vista (IVEE) 

IVEE_2017 <- size_abun_case %>% 
  filter(YEAR == 2017, SITE == "IVEE")

mean(IVEE_2017$SIZE)

IVEE_2012 <- size_abun_case %>% 
  filter(YEAR == 2012, SITE == "IVEE")

#ftest to compare variances in lobster size at NAPL in 2012 and 2017 
# null: Variances are equal (ratio is equal to 1)
# alternative: Variances are not equal (ratio is not equal to 1)

IVEE_ftest <- var.test(IVEE_2012$SIZE, IVEE_2017$SIZE)

IVEE_ftest

#ftest indicates that variances are equal (p =0.307). Use students t-test to compare sizes at NAPL in 2012 and 2017

IVEE_ttest <- t.test(IVEE_2012$SIZE, IVEE_2017$SIZE, var.equal = TRUE)
  
IVEE_ttest

#Our two sample t-test indicates that there is no significant difference in lobster size at Isla Vista (p=0.059) in 2012 (mean = 66.07) and 2017 (mean = 71.4) 

#Site 5: Naples Reef (NAPL) 

NAPL_2017 <- size_abun_case %>% 
  filter(YEAR == 2017, SITE == "NAPL")

mean(NAPL_2017$SIZE)

NAPL_2012 <- size_abun_case %>% 
  filter(YEAR == 2012, SITE == "NAPL")

#ftest to compare variances in lobster size at NAPL in 2012 and 2017 
# null: Variances are equal (ratio is equal to 1)
# alternative: Variances are not equal (ratio is not equal to 1)

NAPL_ftest <- var.test(NAPL_2012$SIZE, NAPL_2017$SIZE)


NAPL_ftest

#ftest indicates that variances are equal (p =0.76). Use students t-test to compare sizes at NAPL in 2012 and 2017

NAPL_ttest <- t.test(NAPL_2012$SIZE, NAPL_2017$SIZE, var.equal = TRUE)
 
 
NAPL_ttest

#Our two sample t-test indicates that there is no significant difference in lobster size at Naples Reef (p=0.5002) in 2012 (mean = 73) and 2017 (mean = 76.23) 



# The only location that differed signficantly in lobster size from 2012-2017 was Mohawk (p= 0.000067). Every other location had no significant difference in lobster sizes from 2012-2017

```

When analzying changes in lobster size, we found that there was no signficant difference in size between 2012 and 2017 at any of the sites, with the exception of Mohawk Reef (t(259) = 4.06, p = 0.000063, $\alpha$ = 0.05). Effect size was medium (Cohen's d = 0.54) and the average size in 2012 was 77.25 mm, compared with 2017 at 72 mm, making the absolute difference in means only 5.25 mm. Two of the five locations (Isla Vista and Naples Reef) are marine protected areas (MPAs), and the other three (Mohawk Reef, Arroyo Quemado, and Carpinteria) are not. Based on our results it appears that there is no relationship between marine protected areas and changes in lobster sizes from 2012-2017.  

### 4. Proportions of “legal” lobsters at the 5 sites in 2017
The legal minimum carapace size for lobster is 82.6 mm. What proportion of observed lobsters at each site are above the legal minimum? Does that proportion differ significantly across the 5 sites? 

```{r}
minimum_carapace <- mean_sizes %>% 
  mutate( 
    carapace = case_when(
      SIZE > 82.6 ~ "Above",
      SIZE < 82.6 ~ "Below"
    )
    ) %>% 
  count(SITE, carapace) %>% 
  spread(carapace, n) %>% 
  select(-SITE)

rownames(minimum_carapace) <- c("AQUE", "CARP","IVEE","MOHK","NAPL")
minimum_carapace


```
```{r}
# Actual proportions:

carapace_prop <- prop.table(as.matrix(minimum_carapace), 1)
carapace_prop

# Chi-square test for independence
carapace_x2 <- chisq.test(minimum_carapace)
carapace_x2
```

2017 proportions of lobster carapace sizes that were above the legal minimum of 82.6 mm differed signficantly by site (x^2^(4)= 18.49, p = 0.0009, $\alpha$ = 0.05). At all sites, the majority of lobster carapace sizes are below the legal minimum. Mohawk Reef had the largest proportion of lobsters below the legal minimum (86%), while Naples Reef had the smallest (66%). This is consistent with our earlier findings that Naples Reef had the largest mean lobster size in 2017. Because the majority of lobsters were smaller than the legal minimum catch size, we don't see a signficant increase in lobster size with decrease in fishing pressure from 2012 to 2017. However, Mohawk, the only location where size changed signficantly, has the largest proportion of smaller lobsters in 2017. As fishing pressure decreased in the channel, we would expect the size to increase, however, sizes actually decreased at Mohawk. 


